---
title: "Yijia-Qiao-HW02"
author: "Yijia Qiao"
date: "9/28/2017"
output: github_document
---

##1. Importing data

#### a) Importing with "read.csv()"
```{r}
dat <- read.csv("/Users/yijiaqiao/stat133/stat133-hws-fall17/hw02/data/nba2017-player-statistics.csv",colClasses = c("Player" = "character", "Team" = "character", "Experience" = "character", "Salary" = "double"))
str(dat)
```

#### b) Importing with "read_csv()"
```{r}
library(readr)
dat2 <- read_csv("/Users/yijiaqiao/stat133/stat133-hws-fall17/hw02/data/nba2017-player-statistics.csv", col_types = cols( Player = "c", Team = "c", Position = col_factor(levels = c('C', 'PF', 'PG', 'SF', 'SG'))) )
str(dat2)
```

##2. Right after importing data
```{r}
dat$Experience[dat$Experience == "R"] <- 0
dat$Experience <- as.integer(dat$Experience)
class(dat$Experience)
```

##3. Performance of players
```{r}
library(dplyr)
#Compute and Add necessary statistics
Missed_FG <- dat$FGA - dat$FGM
Missed_FT <- dat$FTA - dat$FTM
PTS <- 1 * dat$FTM + 2 * dat$Points2 + 3 * dat$Points3 
REB <- dat$OREB + dat$DREB
MPG <- dat$MIN / dat$GP

#Compute EFF as "Efficiency"
EFF <- (PTS + REB + dat$AST + dat$STL + dat$BLK - Missed_FG - Missed_FT - dat$TO) / dat$GP

#Add EFF to dat
dat3 <- mutate(dat, EFF)
summary(EFF)
hist(EFF, main = "Histogram of Efficiency (EFF)", col = "gray")

#Display the player name, team, salary, and EFF value of the top-10 players by EFF in decreasing order (display this information in a data frame).
dat_top10 <- slice(select(arrange(dat3, desc(EFF)), c('Player', 'Team', 'Salary', 'EFF')), 1:10)
dat_top10

#Provide the names of the players that have a negative EFF.
select(filter(dat3, EFF < 0), Player)

#Use the function cor() to compute the correlation coefficients between EFF and all the variables used in the EFF formula.
r_data <- data.frame(
  r = c("r1" = cor(EFF, y = PTS, use = "everything"),
  "r2" <- cor(EFF, y = REB, use = "everything"),
  cor(EFF, y = dat$STL, use = "everything"),
  cor(EFF, y = dat$AST, use = "everything"),
  cor(EFF, y = dat$BLK, use = "everything"),
  -cor(EFF, y = Missed_FG, use = "everything"),
  -cor(EFF, y = Missed_FT, use = "everything"),
  -cor(EFF, y = dat$TO, use = "everything"))
  )
r_data <- arrange(r_data,desc(r))

#Barplot the Correlation
barplot(r_data$r, main = "Correlations between Player Stats and EFF", names.arg = c("PTS", "REB", "STL", "AST", "BLK", "Missed_FT", "Missed_FG", "TO"), axes = TRUE, cex.axis = par("cex.axis"), col = c("lightgray", "lightgray", "lightgray", "lightgray", "lightgray", "mistyrose", "mistyrose", "mistyrose"), ylim = c(-1,1), cex.names = 0.5, border = NA)
abline(h = 0)
```

##4. Efficiency and Salary
```{r}
#Scatterplot between Efficiency and Salary
plot(EFF, dat$Salary, main = "Scatterplot between Efficiency and Salary", xlab = "EFF(Efficiency)", ylab = "Salary", col = "#F15C3C55", pch = 19, las = 1, cex = 0.9)

lines(lowess(EFF, dat$Salary), lwd = 3, col = "blue")
text(25, 16000000, labels = "lowess", col = "blue", cex = 1.5)

#Compute correlation between Efficiency and Salary
cor(EFF, dat$Salary, use = "everything")

#Comments on r between EFF and Salary: as the correlation coefficient is about 0.655, it means there is a positive correlation between salary and efficiency. The Salary increases as the efficiency increases.

```
```{r}
#Rookie players
dat4 <- mutate(dat3, MPG)
players2 <- filter(dat4, MPG >= 20)
players2

#Scatterplot
plot(players2$EFF, players2$Salary, main = "Scatterplot between Efficiency and Salary for Rookie Players", xlab = "EFF(Efficiency)", ylab = "Salary", col = "#F15C3C55", pch = 19, las = 1, cex = 0.9)

lines(lowess(players2$EFF, players2$Salary), lwd = 3, col = "blue")
text(25, 16000000, labels = "lowess", col = "blue", cex = 1.5)

#Compute correlation between Efficiency and Salary
cor(players2$EFF, players2$Salary, use = "everything")

#Comments on r between EFF and Salary for Rookie Players: as the correlation coefficient is about 0.537, it means there is a positive correlation between salary and efficiency. The salary increases as the efficiency inc vvreases.
```

##5. Comments and Reflections
####Reflect on what was hard/easy, problems you solved, helpful tutorials you read, etc.
#####What things were hard, even though you saw them in class/lab?
* I was struggled with displaying x-axis on the bar chart of correlation.

#####What was easy(-ish) even though we haven???t done it in class/lab?
* Calculating correlation factors.

#####Did you need help to complete the assignment? If so, what kind of help?
* Yes, from lab assistants.

#####How much time did it take to complete this HW?
* About 2 hours.

#####What was the most time consuming part?
* Importing data and writing data dictionary.

#####Was there anything that you did not understand? or fully grasped?
* None.

#####Was there anything frustrating in particular?
* Importing data part and converting into required data types.

#####Was there anything exciting? Something that you feel proud of?
* Solving all problems by myself is so exciting.