---
title: "HW01"
author: "Yijia Qiao"
date: "9/22/2017"
output: github_document
---
##1. Importing data
```{r, eval = TRUE}
# load data file
load("/Users/yijiaqiao/stat133/stat133-hws-fall17/hw01/data/nba2017-salary-points.RData")

# list the available objects
ls()
```

##2. Research Questions

###1) A bit of data processing
```{r, eval = TRUE}
# data processing
millions <- round (salary / 1000000 , digits = 2)

experience[experience == "R"] <- 0

new_experience <- as.integer(experience)

new_position <- position

new_position[new_position == 'C'] <- 'center'
new_position[new_position == 'SF'] <- "small_fwd"
new_position[new_position == 'PF'] <- "power_fwd"
new_position[new_position == 'SG'] <- "shoot_guard"
new_position[new_position == 'PG'] <- "point_guard"

new_position <- factor(new_position)

# compute the frequencies of new_position
table(new_position)
```

###2) Scatterplot of points and salary
```{r, eval = TRUE}
# scatterplot of points and salary
plot(points, millions, xlab = "Points", ylab = 'Salary(in miilions)', main = "Scatterplot of Points and Salary", col = 'red')
```

###3) Correlation between points and salary
```{r, eval = TRUE}
# Correlation between points and salary
n <- length(points)
mean_points <- sum(points) / n
mean_salary <- sum(millions) / n
var_points <- 1/(n-1) * sum((points - mean_points)^2)
var_salary <- 1/(n-1) * sum((millions - mean_salary)^2)
sd_points <- sqrt(var_points)
sd_salary <- sqrt(var_salary)
cov_points_salary <- 1/(n-1) * sum((points - mean_points) * (millions - mean_salary))
cor_points_salary <- cov_points_salary / (sd_salary * sd_points)
cor_points_salary
```

###4) Simple linear regression
```{r, eval = TRUE}
# simple linear regression
b1 <- cor_points_salary * sd_salary / sd_points
b0 <- mean_salary - b1 * mean_points
y_hat <- b0 + b1 * points
summary(y_hat)
```

####Answer questions:
#####**What is the regression equation?**
$\hat{Y} = b_0 + b_1 X$
y_hat = `r b0` + `r b1` * X

#####**How do you interpret b1?**
The slope explains how salary relates to points that one player earned. As b1 is positive, it means that when one gets more points, the more he/she earns. 

#####**How do you interpret b0?**
The intercept for salary is an error for predicting. It corrects our predictions based only on relation between points and salary. 

#####**What are predicted salaries for the following?**
```{r, eval = TRUE}
salary_0 <- b0 + b1 * 0
salary_100 <- b0 + b1 * 100
salary_500 <- b0 + b1 * 500
salary_1000 <- b0 + b1 * 1000
salary_2000 <- b0 + b1 * 2000

```

###5) Plotting the regression line
```{r, eval = TRUE}
plot(points, millions, xlab = "Points", ylab = 'Salary(in miilions)', main = "Scatterplot with regression", col = 'yellow')
abline(a = b0, b = b1, col = "blueviolet", lwd = 2 )
lines(lowess(points, millions), col = "red", lwd = 2)
text(2000, 15, "regression", col = "blueviolet")
text (2000, 25, "lowess", col = "red")
```

###6) Regression  residuals and Coefficient of Determination R^2
```{r, eval = TRUE}
e_i <- millions - y_hat
summary(e_i)
RSS <- sum(e_i ^ 2)
TSS <- sum((millions - mean_salary)^2)
R_square <- 1 - RSS/TSS
```

###7) Exploring Position and Experience
```{r, eval = TRUE}
plot(new_experience, millions, col = "deeppink", main = "Scatterplot with lowess smooth")
lines(lowess(new_experience, millions), col = "gray7", lwd = 2)
library("scatterplot3d")
scatterplot3d(points, new_experience, millions, main = "3D scatterplot", color = "darksalmon")

boxplot(millions ~ new_position, main = "Boxplot", ylab = "Salary (in millions)", xlab = "Position")
```

####Answer questions:
#####**Concise descriptions:**
Although the scatterplot and 3d scatterplot shows in different dimensions, they all reveal the fact that if the player stays longer in the team, he/she gets more salaries. As for the boxplot, even players have various positions, their salaries still have a strong relation with their years in experience.

#####**From the scatterplot, does Experience seem to be related with Salary?**
The relation line is rarely smooth and nearly straight. It seems that salary is slightly related to experience.

#####**From the boxplot, does Position seem to be related with Salary?**
Since five boxes have highly similar distribution patterns, it seems that Position is not related with Salary.

###8) Comments and Reflections
#####**What things were hard, even though you saw them in class?**
As for me, how to subset boxplot into 5 boxes is relatively hard comparing with other questions in this homework.
  
#####**What was easy(-ish) even though we haven???t done it in class?**
Labeling plots and changing plot patterns or colors are easy.

#####**If this was the first time you were using git, how do you feel about it?**
It's a little complicated.

#####**If this was the first time using GitHub, how do you feel about it?**
The output file and the way it restored is really simple and easy to use.

#####**Did you need help to complete the assignment? If so, what kind of help? Who helped you?**
I asked lab assistant and GSi for help with 3d scatterplot and boxplot syntaxes.

#####**How much time did it take to complete this HW?**
About two hours.

#####**What was the most time consuming part?**
For 3d scatterplot part.

#####**Was there anything that you did not understand? or fully grasped?**
I did not understand the way that lowess line compiuted.

#####**Was there anything frustrating in particular?**
The most frustrated part is that even I installed 3d packages and run successfully in the console, I still could not knit it.

#####**Was there anything exciting? Something that you feel proud of?**
I am so proud that I finished the first homework.