---
title: "up02-Yijia-Qiao"
author: "Yijia Qiao"
date: "9/15/2017"
output: html_document
---
#**1. Data Analysis for Quantitative Variables** 

``` {r}
load("/Users/yijiaqiao/stat133/stat133-hws-fall17/warmup02/data/nba2017-salary-points.RData")
ls()
```

``` {r}
summary(points)
mean(points)
sd(points)
min(points)
max(points)
median(points)
quantile(points)
IQR(points)
boxplot(points)
# The histogram along with the density curve of points.
hist(points, prob = TRUE, col = "lavender")
lines(density(points), col = "blue")
```

#**2. Data Analysis for Qualitative Variables** 
```{r}
factor (team)
A <- table (team)
B <- A/length(team)
barplot(A, col = "lavender")
title(main = "frequency table of team", font.main = 6)

barplot(B, col = 'mistyrose')
title(main = "relative frequency of team", font.main = 6)
```

#**3. Warmup02 Summary** 
###What is the overall shape?
* The overall shape of the histogram is right-tailed.

###What is the typical center?
* The typical center of the histogram is the median that divides the diagram into two parts of even area.

###Overall range, along with an interval of typical measurements.
* The players score points from 0 to 2558. The first quartile is 156 points and the third quartile is 780 points. The median is 432 points and all scores average at 546.6 points. 